\documentclass[11pt]{article}

\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[margin=1.0in]{geometry}
\usepackage[none]{hyphenat}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{fancyhdr}

\thispagestyle{fancy}
\headheight14pt
\fancyhead{}
\fancyhead[L]{\slshape\MakeUppercase{\large Assignment-04}}
\fancyhead[R]{\large \slshape Victor Solis | \today}

\begin{document}
\begin{flushleft}
Assignment 4: Planning

\section*{Question 1}

Consider the following Crypt-arithmetic problems, where all letters represent a
different digit and the resulting sum is correct. \textbf{Write out all
variables, domains and constraints of the problem.}

    \begin{enumerate}[label=(\alph*)]
        \itemsep1em
        \item SATURN + URANUS = PLANETS
        \item YES + SEND + ME + MORE = MONEY
    \end{enumerate}

\rule[0.1pt]{40em}{1.0pt}

\section*{Question 2}
Consider the following set of edges between nodes. Find a coloring using colors
red, blue, and green such that no two adjacent nodes are assigned the same
color.

\begin{center}
$\{ (a, b),(a, d),(b, c),(b, d),(b, g),(c, g),(d, e),(d, f),(d, g),(f, g) \}$
\end{center}

\begin{enumerate}[label=(\alph*)]
\itemsep1em

    \item Define a CSP for this problem. Clearly define the variables, domains,
        and constraints.
        
    \item Draw the binary constraint graph for this CSP.

    \item Find at least one solution to the CSP. 

\end{enumerate}
\rule[0.1pt]{40em}{1.0pt}

\section*{Question 3}
Consider a block stacking robot with the following actions:

\begin{itemize}
    \item[$\blacksquare$] Stack(x, y)
        \begin{itemize}
            \item Preconditions: Clear(y), Holding(x)
            \item Effects: armEmpty, On(x, y), ¬Clear(y), ¬Holding(x)
        \end{itemize}
    \item[$\blacksquare$] Unstack(x, y)
        \begin{itemize}
            \item Preconditions: Clear(x), On(x, y), armEmpty
            \item Effects: ¬armEmpty, ¬On(x, y), Clear(y), Holding(x)
        \end{itemize}
    \item[$\blacksquare$] Pickup(x)
        \begin{itemize}
            \item Preconditions: Clear(x), On(x, TABLE), armEmpty
            \item Effects: ¬armEmpty, ¬On(x, TABLE), Holding(x)
        \end{itemize}
    \item[$\blacksquare$] Putdown(x)
        \begin{itemize}
            \item Preconditions: Holding(x)
            \item Effects: armEmpty, On(x, TABLE), ¬Holding(x)
        \end{itemize}
\end{itemize}

Create a plan for each of the initial state/goal pairs below
Assume armEmpty is in initial state and the table has infinite space

\begin{enumerate}[label=(\alph*)]
\itemsep1em
    \item TODO
    \item TODO
    \item TODO
\end{enumerate}

\rule[0.1pt]{40em}{1.0pt}

\section*{Question 4}
Consider the following simple planning problem in which the objective is to
interchange the values of two variables v1 and v2

\begin{itemize}
    \item Initial State: Value(v1, 3), Value(v2, 5), Value(v3, 0)
    \item Goal State: Value(v1, 5), Value(v2, 3)
    \item Actions:
        \begin{itemize}
            \item Assign(V, W, X, Y)
                \begin{itemize}
                    \item[$\blacksquare$] Preconditions: Value(V, X),
                        Value(W, Y)
                    \item[$\blacksquare$] Effects: Value(V,
                        Y), ¬Value(W, X)
                \end{itemize}
        \end{itemize}
\end{itemize}


\end{flushleft}
\end{document}
